### 2020-04-04 paper/main.tex
Add paper files

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Transactions Fraud Detection using Machine Learning and Nature Inspired Algorithms}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Peter Mačinec \and Timotej Zaťko}
%
\authorrunning{Peter Mačinec, Timotej Zaťko}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Faculty of Informatics and Information Technologies,\\Slovak University of Technology, Bratislava}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
The abstract should briefly summarize the contents of the paper in
15--250 words.

\keywords{Transactions Fraud Detection \and Machine Learning \and Nature Inspired Algorithms \and Data Analysis.}
\end{abstract}
%
%
%
\section{First Section}
\subsection{A Subsection Sample}
Please note that the first paragraph of a section or subsection is
not indented. The first paragraph that follows a table, figure,
equation etc. does not need an indent, either.

Subsequent paragraphs, however, are indented.

\subsubsection{Sample Heading (Third Level)} Only two levels of
headings should be numbered. Lower level headings remain unnumbered;
they are formatted as run-in headings.

\paragraph{Sample Heading (Fourth Level)}
The contribution should contain no more than four levels of
headings. Table~\ref{tab1} gives a summary of all heading levels.

\begin{table}
\caption{Table captions should be placed above the
tables.}\label{tab1}
\begin{tabular}{|l|l|l|}
\hline
Heading level &  Example & Font size and style\\
\hline
Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
\hline
\end{tabular}
\end{table}

\noindent Displayed equations are centered and set on a separate
line.
\begin{equation}
x + y = z
\end{equation}
Please try to avoid rasterized images for line-art diagrams and
schemas. Whenever possible, use vector graphics instead (see
Fig.~\ref{fig1}).

\begin{figure}
\includegraphics[width=\textwidth]{fig1.eps}
\caption{A figure caption is always placed below the illustration.
Please note that short captions are centered, while long ones are
justified by the macro package automatically.} \label{fig1}
\end{figure}

\begin{theorem}
This is a sample theorem. The run-in heading is set in bold, while
the following text appears in italics. Definitions, lemmas,
propositions, and corollaries are styled the same way.
\end{theorem}
%
% the environments 'definition', 'lemma', 'proposition', 'corollary',
% 'remark', and 'example' are defined in the LLNCS documentclass as well.
%
\begin{proof}
Proofs, examples, and remarks have the initial word in italics,
while the following text appears in normal font.
\end{proof}
For citations of references, we prefer the use of square brackets
and consecutive numbers. Citations using labels or the author/year
convention are also acceptable. The following bibliography provides
a sample reference list with entries for journal
articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs1}, a
book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
and a homepage~\cite{ref_url1}. Multiple citations are grouped
\cite{ref_article1,ref_lncs1,ref_book1},
\cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\begin{thebibliography}{8}
\bibitem{ref_article1}
Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)

\bibitem{ref_lncs1}
Author, F., Author, S.: Title of a proceedings paper. In: Editor,
F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
Springer, Heidelberg (2016). \doi{10.10007/1234567890}

\bibitem{ref_book1}
Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
Location (1999)

\bibitem{ref_proc1}
Author, A.-B.: Contribution title. In: 9th International Proceedings
on Proceedings, pp. 1--2. Publisher, Location (2010)

\bibitem{ref_url1}
LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
Oct 2017
\end{thebibliography}
\end{document}

### 2020-04-04 src/dataset.py
Add option to select a fraction of the dataset

@@ -71,9 +71,11 @@ def split_and_save_processed_data(new_df_transactions, **kwargs):
     y_test.to_csv('../data/y_test.csv', index=False)
 
 
-def load_processed_data():
+def load_processed_data(**kwargs):
     """
     Load preprocessed train-test splitted data.
+
+    :param frac: fraction of axis items to return.
 
     :return: x_train, y_train, x_test, y_test data.
     """
@@ -81,6 +83,14 @@ def load_processed_data():
     y_train = pd.read_csv('../data/y_train.csv')
     x_test = pd.read_csv('../data/x_test.csv')
     y_test = pd.read_csv('../data/y_test.csv')
+    
+    frac = kwargs.get('frac', 1)
+    
+    if frac < 1:
+        x_train.sample(frac=frac, replace=True, random_state=420)
+        y_train.sample(frac=frac, replace=True, random_state=420)
+        x_test.sample(frac=frac, replace=True, random_state=420)
+        y_test.sample(frac=frac, replace=True, random_state=420)        
 
     y_train = y_train['isFraud'].tolist()
     y_test = y_test['isFraud'].tolist()

### 2020-04-04 src/dataset.py
Print number of records

@@ -87,12 +87,21 @@ def load_processed_data(**kwargs):
     frac = kwargs.get('frac', 1)
     
     if frac < 1:
-        x_train.sample(frac=frac, replace=True, random_state=420)
-        y_train.sample(frac=frac, replace=True, random_state=420)
-        x_test.sample(frac=frac, replace=True, random_state=420)
-        y_test.sample(frac=frac, replace=True, random_state=420)        
+        n_train = int(len(x_train) * frac)
+        n_test = int(len(x_test) * frac)
+        
+        x_train = x_train.sample(n=n_train, replace=False, random_state=420)
+        y_train = y_train.sample(n=n_train, replace=False, random_state=420)
+        x_test = x_test.sample(n=n_test, replace=False, random_state=420)
+        y_test = y_test.sample(n=n_test, replace=False, random_state=420)        
 
     y_train = y_train['isFraud'].tolist()
     y_test = y_test['isFraud'].tolist()
+    
+    print(f'Number of records:')
+    print(f'  x_train - {len(x_train)}')
+    print(f'  y_train - {len(y_train)}')
+    print(f'  x_test - {len(x_test)}')
+    print(f'  y_test - {len(y_test)}')
 
     return x_train, y_train, x_test, y_test

### 2020-04-04 src/preprocessing/transformers.py
Add Normalizer transformer

@@ -2,6 +2,7 @@ import pandas as pd
 import numpy as np
 
 from sklearn.base import BaseEstimator, TransformerMixin
+from sklearn.preprocessing import normalize
 
 
 class SelectFeatures(BaseEstimator, TransformerMixin):
@@ -60,12 +61,57 @@ class EmailProviderTransform(TransformerMixin, BaseEstimator):
         return self
 
     def transform(self, x, y=None, copy=None):
-        def transform_email(val):
-            if val is np.nan:
-                return val
-            return val.split('.')[0]
+        def transform_email(value):
+            """
+            Transform email domains into domain names only.
 
-        for column in self.columns:
-            x[column] = x[column].astype('str').apply(transform_email).astype('str')
+            We have found in the data, that some email domains are
+            repeated (e.g. `gmail` and `gmail.com`). Those domains
+            should be aggregated into simply `gmail`.
+
+            :param value: value to be transformed.
+            :return: transformed email domain.
+            """
+            if value is np.nan:
+                return value
+            return value.split('.')[0]
+
+        for col in self.columns:
+            x[col] = x[col].astype('str').apply(transform_email).astype('str')
 
         return x
+
+
+class Normalizer(TransformerMixin):
+    """Normalize numerical attributes of dataframe."""
+
+    def fit(self, X, y=None):
+        return self
+
+    def transform(self, df):
+        df = (df - df.mean()) / df.std()
+        return df
+
+
+class OutliersFilter(TransformerMixin):
+    """Remove outliers using quartiles."""
+
+    def __init__(self):
+        self.Q1 = {}
+        self.Q3 = {}
+
+    def fit(self, df, y=None, **fit_params):
+        for col in df.columns:
+            self.Q1[col] = df[col].quantile(.25)
+            self.Q3[col] = df[col].quantile(.75)
+        return self
+
+    def transform(self, df, **transform_params):
+        for col in df.columns:
+            Q1 = self.Q1[col]
+            Q3 = self.Q3[col]
+            lower_outlier = Q1 - (Q3 - Q1) * 1.5
+            upper_outlier = Q3 + (Q3 - Q1) * 1.5
+
+            df = df[(df[col] >= lower_outlier) & (df[col] <= upper_outlier)]
+        return df

### 2020-04-04 src/dataset.py
Fix loading data with custom types

@@ -24,26 +24,26 @@ def load_data():
     for col in ['isFraud']:
         dtype[col] = np.bool
 
-    for v in range(1, 339):
+    for v in range(1, 340):
         dtype[f'V{v}'] = np.float32
 
-    for d in range(1, 15):
+    for d in range(1, 16):
         dtype[f'D{d}'] = np.float16
 
-    for c in range(1, 14):
+    for c in range(1, 15):
         dtype[f'C{c}'] = np.float16
 
     for i in ['id_01', 'id_02', 'id_03', 'id_04', 'id_05', 'id_06',
               'id_07', 'id_08', 'id_09', 'id_10', 'id_11']:
-        dtype[f'id_{i}'] = np.float32
+        dtype[i] = np.float32
 
-    for c in range(1, 6):
+    for c in range(1, 7):
         dtype[f'card{c}'] = np.object
 
-    for i in range(12, 38):
+    for i in range(12, 39):
         dtype[f'id_{i}'] = np.object
 
-    for a in range(1, 2):
+    for a in range(1, 3):
         dtype[f'addr{a}'] = np.object
 
     return pd.read_csv('../data/dataset.csv', dtype=dtype)

### 2020-04-04 src/preprocessing/transformers.py
Refactor transformers

@@ -2,35 +2,46 @@ import pandas as pd
 import numpy as np
 
 from sklearn.base import BaseEstimator, TransformerMixin
-from sklearn.preprocessing import normalize
 
 
 class SelectFeatures(BaseEstimator, TransformerMixin):
+    """
+    Select subset of features from dataframe.
+    
+    :param columns: columns to choose from dataframe.
+    """
+
     def __init__(self, columns):
         self.columns = columns
 
     def fit(self, x, y=None):
         return self
 
-    def transform(self, x):
-        assert isinstance(x, pd.DataFrame)
+    def transform(self, df):
+        assert isinstance(df, pd.DataFrame)
 
-        return x[self.columns]
+        return df[self.columns]
 
 
 class FilterColumnsByCountOfMissingValues(BaseEstimator, TransformerMixin):
+    """
+    Filter columns by count of missing values.
+    
+    :param threshold: threshold ratio of missing values in column.
+    """
+
     def __init__(self, threshold, **kwargs):
         self.threshold = threshold
 
     def fit(self, x, y=None):
         return self
 
-    def transform(self, x):
-        assert isinstance(x, pd.DataFrame)
+    def transform(self, df):
+        assert isinstance(df, pd.DataFrame)
 
-        columns_to_drop = x.columns[1 - x.isnull().mean() < self.threshold].tolist()
+        to_drop = df.columns[1 - df.isnull().mean() < self.threshold].tolist()
 
-        return x.drop(columns=columns_to_drop)
+        return df.drop(columns=to_drop)
 
 
 class KeepOnlyMostCommonValues(BaseEstimator, TransformerMixin):
@@ -54,13 +65,19 @@ class KeepOnlyMostCommonValues(BaseEstimator, TransformerMixin):
 
 
 class EmailProviderTransform(TransformerMixin, BaseEstimator):
+    """
+    Transform email addresses into domains only.
+
+    :param columns: columns containing email addresses.
+    """
+
     def __init__(self, columns):
         self.columns = columns
 
     def fit(self, x, y=None):
         return self
 
-    def transform(self, x, y=None, copy=None):
+    def transform(self, df, y=None, copy=None):
         def transform_email(value):
             """
             Transform email domains into domain names only.
@@ -77,41 +94,17 @@ class EmailProviderTransform(TransformerMixin, BaseEstimator):
             return value.split('.')[0]
 
         for col in self.columns:
[…]

### 2020-04-04 src/preprocessing/transformers.py
Add transformer to merge small categories

@@ -44,24 +44,48 @@ class FilterColumnsByCountOfMissingValues(BaseEstimator, TransformerMixin):
         return df.drop(columns=to_drop)
 
 
-class KeepOnlyMostCommonValues(BaseEstimator, TransformerMixin):
-    def __init__(self, n, **kwargs):
-        self.n = n
-        self.columns = kwargs.get('columns', None)
+class MergeSmallCategories(TransformerMixin):
+    """
+    Merge too small classes in categorical attribute into 'other' class.
 
-    def fit(self, x, y=None):
+    Taken from other author's project:
+    https://github.com/pmacinec/diabetes-patients-readmissions-prediction/
+
+    :param threshold: threshold of percentage frequency under which 
+        classes will be merged into 'other' class.
+    """
+
+    def __init__(self, threshold=0.05, **kwargs):
+        self.threshold = threshold
+        self.mapping = {}
+
+    def fit(self, df, y=None):
+        for col in df.columns:
+            values = df[col].value_counts(normalize=True)
+            for name, value in values.iteritems():
+                if value < self.threshold:
+                    self.mapping[name] = 'other'
         return self
 
-    def transform(self, x):
-        assert isinstance(x, pd.DataFrame)
+    def transform(self, df, **transform_params):
+        for col in df.columns:
+            df[col] = df[col].apply(
+                lambda x: self.get_value(x)
+            )
+        return df
 
-        columns = self.columns if self.columns is not None else x.columns
+    def get_value(self, value):
+        """
+        Get value from mapping with handling special cases.
 
-        for column in columns:
-            most_common = x[column].value_counts().nlargest(n=self.n).index.tolist()
-            x.loc[~x[column].isin(most_common), column] = 'Other'
-
-        return x[columns]
+        :param value: value to be found in mapping.
+        :return: mapped value or None, if value is unknown or NaN.
+        """
+        if pd.isna(value):
+            return None
+        if value not in self.mapping.keys():
+            return value
+        return self.mapping[value]
 
 
 class EmailProviderTransform(TransformerMixin, BaseEstimator):

### 2020-04-04 src/preprocessing/transformers.py
Fix email transformer for missing column

@@ -118,6 +118,8 @@ class EmailProviderTransform(TransformerMixin, BaseEstimator):
             return value.split('.')[0]
 
         for col in self.columns:
+            if col not in df.columns:
+                continue
             df[col] = df[col].astype('str').apply(transform_email).astype('str')
 
         return df

### 2020-04-04 src/preprocessing/transformers.py
Refactor normalizer transformer

@@ -128,9 +128,21 @@ class EmailProviderTransform(TransformerMixin, BaseEstimator):
 class Normalizer(TransformerMixin):
     """Normalize numerical attributes of dataframe."""
 
+    def __init__(self):
+        self.means = {}
+        self.stds = {}
+
     def fit(self, df, y=None):
+        for col in df.columns:
+            if '_missing' in col:
+                continue
+            self.means[col] = df[col].astype(float).mean()
+            self.stds[col] = df[col].astype(float).std()
         return self
 
     def transform(self, df):
-        df = (df - df.mean()) / df.std()
+        for col in df.columns:
+            if '_missing' in col:
+                continue
+            df[col] = (df[col] - self.means[col]) / self.stds[col]
         return df

### 2020-04-04 src/preprocessing/pandas_feature_union.py
Refactor some transformers

@@ -9,6 +9,8 @@ from scipy import sparse
 
 
 class PandasFeatureUnion(FeatureUnion):
+    """Feature union transformer for pandas."""
+
     def fit_transform(self, X, y=None, **fit_params):
         self._validate_transformers()
         result = Parallel(n_jobs=self.n_jobs)(

### 2020-04-04 src/preprocessing/pandas_missing_indicator.py

@@ -5,10 +5,12 @@ from sklearn.impute import MissingIndicator
 
 
 class PandasMissingIndicator(MissingIndicator):
+    """Missing indicator transformer for pandas."""
+
     def __init__(self, **kwargs):
         super().__init__(features='all', **kwargs)
 
-        self._suffix = kwargs.get('suffix', ' missing')
+        self._suffix = kwargs.get('suffix', '_missing')
 
     def fit(self, x, y=None):
         super().fit(x, y)

### 2020-04-04 src/preprocessing/pandas_one_hot_encoder.py

@@ -4,10 +4,11 @@ from sklearn.base import BaseEstimator, TransformerMixin
 
 
 class PandasOneHotEncoder(BaseEstimator, TransformerMixin):
-    def fit(self, x, y=None):
+    def fit(self, df, y=None):
         return self
 
-    def transform(self, x):
-        assert isinstance(x, pd.DataFrame)
+    def transform(self, df):
+        assert isinstance(df, pd.DataFrame)
 
-        return pd.get_dummies(x, dummy_na=True, drop_first=True)
+        columns = list(filter(lambda x: '_missing' not in x, df.columns))
+        return pd.get_dummies(df, columns=columns)

### 2020-04-04 src/preprocessing/pandas_simple_imputer.py

@@ -4,6 +4,8 @@ from sklearn.impute import SimpleImputer
 
 
 class PandasSimpleImputer(SimpleImputer):
+    """Simple imputer transformer for pandas."""
+    
     def __init__(self, **kwargs):
         super().__init__(**kwargs)
 

### 2020-04-04 src/preprocessing/transformers.py

@@ -61,6 +61,8 @@ class MergeSmallCategories(TransformerMixin):
 
     def fit(self, df, y=None):
         for col in df.columns:
+            if '_missing' in col:
+                continue
             values = df[col].value_counts(normalize=True)
             for name, value in values.iteritems():
                 if value < self.threshold:
@@ -69,6 +71,8 @@ class MergeSmallCategories(TransformerMixin):
 
     def transform(self, df, **transform_params):
         for col in df.columns:
+            if '_missing' in col:
+                continue
             df[col] = df[col].apply(
                 lambda x: self.get_value(x)
             )

### 2020-04-04 src/preprocessing/pandas_transform.py
Remove unused preprocessing transformer

@@ -1,14 +1,0 @@
-# https://github.com/marrrcin/pandas-feature-union/blob/master/pandas_transform.py
-
-from sklearn.base import BaseEstimator, TransformerMixin
-
-
-class PandasTransform(TransformerMixin, BaseEstimator):
-    def __init__(self, fn):
-        self.fn = fn
-
-    def fit(self, X, y=None):
-        return self
-
-    def transform(self, X, y=None, copy=None):
-        return self.fn(X)

### 2020-04-04 src/preprocessing/pandas_missing_indicator.py
Fix type conversions of missing indicators

@@ -31,6 +31,6 @@ class PandasMissingIndicator(MissingIndicator):
             if np.any(row == True):
                 x.loc[:, col_name] = row
 
-                x[col_name] = x[col_name].astype(np.bool)
+                x[col_name] = x[col_name].astype(int)
 
         return x

### 2020-04-04 src/preprocessing/pandas_simple_imputer.py

@@ -21,6 +21,8 @@ class PandasSimpleImputer(SimpleImputer):
         matrix = super().transform(x)
 
         for index, col in enumerate(x.columns):
+            if '_missing' in col:
+                continue
             x.loc[:, col] = matrix[:, index]
 
         return x

### 2020-04-04 src/evaluation.py
Add ROC metric

@@ -1,4 +1,4 @@
-from sklearn.metrics import classification_report, plot_confusion_matrix
+from sklearn.metrics import classification_report, plot_confusion_matrix, roc_auc_score
 
 
 def custom_classification_report(clf, x_test, y_test):
@@ -29,5 +29,10 @@ def custom_classification_report(clf, x_test, y_test):
             print(f'{key}: {clf_report[key]}')
 
         print('\n')
+
+    print(f'\033[1mArea Under the Receiver Operating Characteristic Curve (ROC AUC)\033[0m')
+    for average in ['micro', 'macro', 'samples', 'weighted']:
+        print(f'{average}: {roc_auc_score(y_test, y_pred, average=average)}')
+    print('\n')
 
     plot_confusion_matrix(clf, x_test, y_test)

### 2020-04-04 src/classification_benchmark.py
Use GWO to find optimal features

@@ -1,0 +1,61 @@
+from NiaPy.benchmarks import Benchmark
+
+
+class ClassificationBenchmark(Benchmark):
+    """
+    NiaPy benchmark for classification task
+
+    :param model_fn: function which returns sklearn model
+    :param eval_fn: function(y_test, y_pred) which evaluates predictions and returns a scalar
+    :param x_train: train data
+    :param y_train: train labels
+    :param x_test: test data
+    :param y_test: test labels
+    """
+    def __init__(self, model_fn, eval_fn, x_train, y_train, x_test, y_test):
+        self.x_train = x_train
+        self.y_train = y_train
+        self.x_test = x_test
+        self.y_test = y_test
+        self.model_fn = model_fn
+        self.eval_fn = eval_fn
+
+        self.columns = x_train.columns
+
+        Benchmark.__init__(self, 0, 1)
+
+    def get_length(self):
+        """
+        Get length of the vector which is being optimized.
+        :return: length of the vector which is being optimized
+        """
+        return len(self.columns)
+
+    def select_columns(self, solution_vec):
+        """
+        Select columns based on the solution vector.
+        :param solution_vec: current solution of the problem as a vector
+        :return: list of column names based on the solution vector
+        """
+        return list(self.columns[solution_vec > 0.5])
+
+    def function(self):
+        def evaluate(_, solution_vec):
+            selected_columns = self.select_columns(solution_vec)
+
+            if len(selected_columns) < 1:
+                # return the inverted score, since the optimizer minimizes the task
+                return 1 - 1
+
+            clf = self.model_fn()
+            clf = clf.fit(self.x_train[selected_columns], self.y_train)
+
+            y_pred = clf.predict(self.x_test[selected_columns])
+            score = self.eval_fn(self.y_test, y_pred)
+
+            # print(len(selected_columns), score)
+
+            # return the inverted score, since the optimizer minimizes the task
+            return 1 - score
+
+        return evaluate

### 2020-04-04 src/dataset.py
Refactor splitting helper

@@ -49,17 +49,17 @@ def load_data():
     return pd.read_csv('../data/dataset.csv', dtype=dtype)
 
 
-def split_and_save_processed_data(new_df_transactions, **kwargs):
+def split_and_save_processed_data(df, **kwargs):
     """
     Split and store dataframe into train-test dataframes.
 
-    :param new_df_transactions: dataframe to be splitted.
+    :param df: dataframe to be splitted.
     :param kwargs: additional arguments:
         test_size (float): test data ratio (default 0.2).
     """
     print('Splitting the data...')
-    x = new_df_transactions.drop('isFraud', axis=1)
-    y = new_df_transactions[['isFraud']]
+    x = df.drop('isFraud', axis=1)
+    y = df[['isFraud']]
     x_train, x_test, y_train, y_test = train_test_split(
         x, y, test_size=kwargs.get('test_size', 0.2), random_state=42
     )

### 2020-04-06 src/algorithms/gray_wolf_optimizer_multiprocessing.py
Add GWO with multiprocessing

@@ -1,0 +1,63 @@
+from pathos.multiprocessing import ProcessPool
+
+from numpy import fabs
+
+from NiaPy.algorithms.basic import GreyWolfOptimizer
+
+
+class GreyWolfOptimizerMultiprocessing(GreyWolfOptimizer):
+	def __init__(self, **kwargs):
+		super().__init__(**kwargs)
+
+		self.nodes = kwargs.get('nodes', 4)
+
+	def runIteration(self, task, pop, fpop, xb, fxb, A, A_f, B, B_f, D, D_f, **dparams):
+		r"""Core funciton of GreyWolfOptimizer algorithm.
+		Args:
+			task (Task): Optimization task.
+			pop (numpy.ndarray): Current population.
+			fpop (numpy.ndarray): Current populations function/fitness values.
+			xb (numpy.ndarray):
+			fxb (float):
+			A (numpy.ndarray):
+			A_f (float):
+			B (numpy.ndarray):
+			B_f (float):
+			D (numpy.ndarray):
+			D_f (float):
+			**dparams (Dict[str, Any]): Additional arguments.
+		Returns:
+			Tuple[numpy.ndarray, numpy.ndarray, numpy.ndarray, float, Dict[str, Any]]:
+				1. New population
+				2. New population fitness/function values
+				3. Additional arguments:
+					* A (): TODO
+		"""
+
+		def eval_task(args):
+			i, w = args
+			A1, C1 = 2 * a * self.rand(task.D) - a, 2 * self.rand(task.D)
+			X1 = A - A1 * fabs(C1 * A - w)
+			A2, C2 = 2 * a * self.rand(task.D) - a, 2 * self.rand(task.D)
+			X2 = B - A2 * fabs(C2 * B - w)
+			A3, C3 = 2 * a * self.rand(task.D) - a, 2 * self.rand(task.D)
+			X3 = D - A3 * fabs(C3 * D - w)
+			pop = task.repair((X1 + X2 + X3) / 3, self.Rand)
+			fpop = task.eval(pop[i])
+			return i, pop, fpop
+
+		a = 2 - task.Evals * (2 / task.nFES)
+		pool = ProcessPool(nodes=4)
+		pool.clear()
+		results = pool.map(eval_task, [[i, w] for i, w in enumerate(pop)])
+
+		for i, _pop, _fpop in results:
+			pop[i] = _pop
[…]

### 2020-04-06 src/classification_benchmark.py

@@ -1,5 +1,7 @@
 from NiaPy.benchmarks import Benchmark
 
+import pandas as pd
+import numpy as np
 
 class ClassificationBenchmark(Benchmark):
     """
@@ -20,8 +22,6 @@ class ClassificationBenchmark(Benchmark):
         self.model_fn = model_fn
         self.eval_fn = eval_fn
 
-        self.columns = x_train.columns
-
         Benchmark.__init__(self, 0, 1)
 
     def get_length(self):
@@ -29,7 +29,7 @@ class ClassificationBenchmark(Benchmark):
         Get length of the vector which is being optimized.
         :return: length of the vector which is being optimized
         """
-        return len(self.columns)
+        return len(self.x_train.columns)
 
     def select_columns(self, solution_vec):
         """
@@ -37,23 +37,25 @@ class ClassificationBenchmark(Benchmark):
         :param solution_vec: current solution of the problem as a vector
         :return: list of column names based on the solution vector
         """
-        return list(self.columns[solution_vec > 0.5])
+        return self.x_train.columns[solution_vec > 0.5].tolist()
 
     def function(self):
         def evaluate(_, solution_vec):
             selected_columns = self.select_columns(solution_vec)
 
+            # fix for incorrect serialization when using multi threading module
+            if len(selected_columns) == 1 and not isinstance(selected_columns[0], str):
+                selected_columns = selected_columns[0]
+
             if len(selected_columns) < 1:
                 # return the inverted score, since the optimizer minimizes the task
-                return 1 - 1
+                return 1 - 0
 
             clf = self.model_fn()
             clf = clf.fit(self.x_train[selected_columns], self.y_train)
 
             y_pred = clf.predict(self.x_test[selected_columns])
             score = self.eval_fn(self.y_test, y_pred)
-
-            # print(len(selected_columns), score)
 
             # return the inverted score, since the optimizer minimizes the task
             return 1 - score

### 2020-04-07 src/algorithms/gray_wolf_optimizer_multiprocessing.py
Fix GWO nodes count

@@ -47,7 +47,7 @@ class GreyWolfOptimizerMultiprocessing(GreyWolfOptimizer):
 			return i, pop, fpop
 
 		a = 2 - task.Evals * (2 / task.nFES)
-		pool = ProcessPool(nodes=4)
+		pool = ProcessPool(nodes=self.nodes)
 		pool.clear()
 		results = pool.map(eval_task, [[i, w] for i, w in enumerate(pop)])
 

### 2020-04-18 src/data_balancing.py
Add data balancing helpers
For undersampling, only random undersampling is used. For oversampling, SMOTE method is used.

@@ -1,0 +1,26 @@
+from imblearn.under_sampling import RandomUnderSampler
+from imblearn.over_sampling import SMOTE
+
+
+def random_undersample(x, y):
+    """
+    Undersample data by random choosing samples from majority class.
+
+    :param x: train data features.
+    :param y: train data labels.
+    :return: x, y after undersampling.
+    """
+    rus = RandomUnderSampler(random_state=42)
+    return rus.fit_resample(x, y)
+
+
+def smote_oversample(x, y):
+    """
+    Oversample data using SMOTE method.
+
+    :param x: train data features.
+    :param y: train data labels.
+    :return: x, y after oversampling.
+    """
+    smote = SMOTE(random_state=42)
+    return smote.fit_resample(x, y)

### 2020-04-18 src/evaluation.py
Simplify ROC AUC score printout

@@ -1,4 +1,5 @@
-from sklearn.metrics import classification_report, plot_confusion_matrix, roc_auc_score
+from sklearn.metrics import classification_report, plot_confusion_matrix,\
+    roc_auc_score
 
 
 def custom_classification_report(clf, x_test, y_test):
@@ -30,9 +31,9 @@ def custom_classification_report(clf, x_test, y_test):
 
         print('\n')
 
-    print(f'\033[1mArea Under the Receiver Operating Characteristic Curve (ROC AUC)\033[0m')
-    for average in ['micro', 'macro', 'samples', 'weighted']:
-        print(f'{average}: {roc_auc_score(y_test, y_pred, average=average)}')
+    print(f'\033[1mArea Under the Receiver Operating Characteristic Curve '
+          f'(ROC AUC)\033[0m')
+    print(f'Score: {roc_auc_score(y_test, y_pred)}')
     print('\n')
 
     plot_confusion_matrix(clf, x_test, y_test)

### 2020-04-18 src/algorithms/grey_wolf_optimizer_multiprocessing.py
Rename GWO multiprocessing file

@@ -1,7 +1,5 @@
 from pathos.multiprocessing import ProcessPool
-
 from numpy import fabs
-
 from NiaPy.algorithms.basic import GreyWolfOptimizer
 
 

### 2020-04-18 paper/main.tex
Update paper structure

@@
 %
-\section{First Section}
-\subsection{A Subsection Sample}
-Please note that the first paragraph of a section or subsection is
-not indented. The first paragraph that follows a table, figure,
-equation etc. does not need an indent, either.
+\section{Introduction}
 
-Subsequent paragraphs, however, are indented.
+% TODO: co robime, tj, nacrtnut problem atd
+% TODO: spomenut, datasety s velkym mnozstvom features, je potrebna feature selection
+% TODO: spomenut, ze robi sa to takto takto, ale su aj nejako prirodou inspirovane algoritmy, ktore maju taketo vyhody
+% TODO: preco? napr. mzoe chybat domenovat znalost (features) kvoli anonymizacii - nas pripad
 
-\subsubsection{Sample Heading (Third Level)} Only two levels of
-headings should be numbered. Lower level headings remain unnumbered;
-they are formatted as run-in headings.
+\section{Related works}
 
-\paragraph{Sample Heading (Fourth Level)}
-The contribution should contain no more than four levels of
-headings. Table~\ref{tab1} gives a summary of all heading levels.
+% TODO: pribuzne prace na feature selection
+% TODO: spomenut pouzite algoritmy
+% TODO: spomenut domeny, datasety, charakteristky datasetov
 
-\begin{table}
-\caption{Table captions should be placed above the
-tables.}\label{tab1}
-\begin{tabular}{|l|l|l|}
-\hline
-Heading level &  Example & Font size and style\\
-\hline
-Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
-1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
-2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
-3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
-4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
-\hline
-\end{tabular}
-\end{table}
+- Gray Wolf Optimization Algorithm \cite{Mirjalili_Mirjalili_Lewis_2014}
 
-\noindent Displayed equations are centered and set on a separate
-line.
-\begin{equation}
-x + y = z
-\end{equation}
-Please try to avoid rasterized images for line-art diagrams and
-schemas. Whenever possible, use vector graphics instead (see
-Fig.~\ref{fig1}).
+\section{Problem definition}
 
-\begin{figure}
-\includegraphics[width=\textwidth]{fig1.eps}
-\caption{A figure caption is always placed below the illustration.
-Please note that short captions are centered, while long ones are
-justified by the macro package automatically.} \label{fig1}
-\end{figure}
+% TODO: opisat nas problem
+% TODO: spomenut - nevyvazene triedy, velmi vela features, chyba domenovat znalost (features) kvoli anonymizacii
+% TODO: opisat dataset - nejaky grafy, pocty pozorovani, features
 
-\begin{theorem}
-This is a sample theorem. The run-in heading is set in bold, while
-the following text appears in italics. Definitions, lemmas,
-propositions, and corollaries are styled the same way.
-\end{theorem}
-%
-% the environments 'definition', 'lemma', 'proposition', 'corollary',
-% 'remark', and 'example' are defined in the LLNCS documentclass as well.
-%
-\begin{proof}
-Proofs, examples, and remarks have the initial word in italics,
-while the following text appears in normal font.
-\end{proof}
-For citations of references, we prefer the use of square brackets
-and consecutive numbers. Citations using labels or the author/year
-convention are also acceptable. The following bibliography provides
-a sample reference list with entries for journal
-articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs1}, a
-book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
-and a homepage~\cite{ref_url1}. Multiple citations are grouped
-\cite{ref_article1,ref_lncs1,ref_book1},
-\cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
-%
-% ---- Bibliography ----
-%
-% BibTeX users should specify bibliography style 'splncs04'.
-% References will then be sorted and formatted in the correct style.
-%
-% \bibliographystyle{splncs04}
-% \bibliography{mybibliography}
-%
-\begin{thebibliography}{8}
-\bibitem{ref_article1}
-Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
+\section{Method proposal}
 
-\bibitem{ref_lncs1}
-Author, F., Author, S.: Title of a proceedings paper. In: Editor,
-F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
-Springer, Heidelberg (2016). \doi{10.10007/1234567890}
+% TODO: len kratko, co navrhujeme - feature selection na nejaky model pomocou nejakeho prirodou inspirovaneho algoritmu
+% TODO: mozeno spomenut oversampling/undersampling
 
-\bibitem{ref_book1}
-Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
-Location (1999)
+\section{Preparation}
 
-\bibitem{ref_proc1}
-Author, A.-B.: Contribution title. In: 9th International Proceedings
-on Proceedings, pp. 1--2. Publisher, Location (2010)
+% TODO: spomenut preprocessing, co sme robili a co z toho vzislo
+% TODO: spomenut transformery, one hot, a filter
+% TODO: kolko bol finalny pocet features
+% TODO: len kratko spomenut ze preco sme sa rozhodli pouzit aky model - decision tree
 
-\bibitem{ref_url1}
-LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
-Oct 2017
-\end{thebibliography}
+\section{Experiments}
+
+% TODO: ake experiemnty sme spravili
+% TODO: tabulka porovnania algoritmov
+% TODO: learning along the way with consequences - napr undersampling/oversampling atd.
+
+\section{Conclusion}
+
+% TODO: zhrnutie vysledkov, mozne vylepsenia a uskalia nasho pristupu
+
+\bibliographystyle{splncs04}
+\bibliography{references}
+
 \end{document}

### 2020-04-18 src/classification_benchmark.py
Refactor classification benchmark class

@@ -1,18 +1,17 @@
 from NiaPy.benchmarks import Benchmark
 
-import pandas as pd
-import numpy as np
 
 class ClassificationBenchmark(Benchmark):
     """
-    NiaPy benchmark for classification task
+    NiaPy benchmark for classification task.
 
-    :param model_fn: function which returns sklearn model
-    :param eval_fn: function(y_test, y_pred) which evaluates predictions and returns a scalar
-    :param x_train: train data
-    :param y_train: train labels
-    :param x_test: test data
-    :param y_test: test labels
+    :param model_fn: function which returns sklearn model.
+    :param eval_fn: function(y_test, y_pred) which evaluates predictions
+         and returns a scalar.
+    :param x_train: train data.
+    :param y_train: train labels.
+    :param x_test: test data.
+    :param y_test: test labels.
     """
     def __init__(self, model_fn, eval_fn, x_train, y_train, x_test, y_test):
         self.x_train = x_train
@@ -27,15 +26,17 @@ class ClassificationBenchmark(Benchmark):
     def get_length(self):
         """
         Get length of the vector which is being optimized.
-        :return: length of the vector which is being optimized
+
+        :return: length of the vector which is being optimized.
         """
         return len(self.x_train.columns)
 
     def select_columns(self, solution_vec):
         """
         Select columns based on the solution vector.
-        :param solution_vec: current solution of the problem as a vector
-        :return: list of column names based on the solution vector
+
+        :param solution_vec: solution of the problem as a vector.
+        :return: list of column names based on the solution vector.
         """
         return self.x_train.columns[solution_vec > 0.5].tolist()
 
@@ -43,12 +44,13 @@ class ClassificationBenchmark(Benchmark):
         def evaluate(_, solution_vec):
             selected_columns = self.select_columns(solution_vec)
 
-            # fix for incorrect serialization when using multi threading module
-            if len(selected_columns) == 1 and not isinstance(selected_columns[0], str):
+            # fix of incorrect serialization when using multi threading module
+            if len(selected_columns) == 1 and \
+                    not isinstance(selected_columns[0], str):
                 selected_columns = selected_columns[0]
 
             if len(selected_columns) < 1:
-                # return the inverted score, since the optimizer minimizes the task
+                # inverted score, since the optimizer minimizes the task
                 return 1 - 0
 
             clf = self.model_fn()
@@ -57,7 +59,7 @@ class ClassificationBenchmark(Benchmark):
[…]

### 2020-04-19 src/evaluation.py
Implement option to disable confussion matrix

@@ -2,15 +2,17 @@ from sklearn.metrics import classification_report, plot_confusion_matrix,\
     roc_auc_score
 
 
-def custom_classification_report(clf, x_test, y_test):
+def custom_classification_report(clf, x_test, y_test, **kwargs):
     """
     Create custom classification report.
 
     :param clf: classifier model.
     :param x_test: test samples to predict labels for.
     :param y_test: true label values of test samples.
+    :param confussion_matrix: wether to plut confussion matrix.
     """
     y_pred = clf.predict(x_test)
+    confussion_matrix = kwargs.get('confussion_matrix', True)
 
     clf_report = classification_report(
         y_pred,
@@ -36,4 +38,5 @@ def custom_classification_report(clf, x_test, y_test):
     print(f'Score: {roc_auc_score(y_test, y_pred)}')
     print('\n')
 
-    plot_confusion_matrix(clf, x_test, y_test)
+    if confussion_matrix:
+        plot_confusion_matrix(clf, x_test, y_test)

### 2020-05-01 paper/main.tex
Add introduction and some parts to paper

@@
 
+Payments with credit cards are nowadays still more preferred over using cash.
+Using credit card is not only more comfortable for people, but even also more safe than carrying cash in wallet when it comes to higher amount of money.
+However, number of transaction frauds is arising alongside with usage of credit cards.
+Number of transactions and ability to obtain data from them indicate the need of automatic detection of suspicious payments.
+
+Data of performed transactions via credit cards are naturally collected by companies and banks to produce statistics or investigate frauds.
+Therefore, transactions fraud detection can be interpreted as data mining problem, concretely binary classification.
+
+In this paper, we propose novel method for detection frauds in transactions using aspects of data analysis, machine learning and nature inspired algorithms.
+The basis of our method lies in training machine learning model on best features selected by nature inspired algorithm.
+More nature inspired algorithms are compared to choose the best one for this problem.
+Finally, nature inspired algorithms in combination with machine learning proved to be very efficient way for detecting frauds in transactions, overperforming common methods in this area.
+% TODO ten overperforming treba nejak doladit mozno
+
+% Required: Úvod / Introduction - aká úloha sa ide riešiť, prečo je dobré ju riešiť
+
 % TODO: co robime, tj, nacrtnut problem atd
@@
 
+% Required: Podobné práce / Related Work - ako riešili vašu alebo podobnú úlohu v
+% existujúcich prácach - nezabudnite v texte uvádzať citácie na práce.
+
+%  opis použitého algoritmu - aj s odôvodnením výberu algoritmu
+%  opis vykonaných experimentov - opis použitého datasetu, opis cieľu a postupu
+% experimentu, opis nastavení algoritmu (odôvodnenie prečo zvolené
+% nastavenia), opis metrík, ktorými experiment vyhodnocujete... ak sa
+% porovnávate s existujúcimi riešeniami, nezabudnite ich citovať, prípadne aj
+% stručne opísať
+%  Vyhodnotenie / Evaluation - prehľadné výsledky experimentov (formou
+% textového opisu, grafov, tabuliek, ...)
+%  Záver / Conclusion
+%  Použitá literatúra / References - používajte LNCS štýl odkazovania sa na
+% zdroje, pokyny nájdete aj v šablóne (príklady tu alebo tu, štýl pre správcu
+% citácii Mendeley)
+
 - Gray Wolf Optimization Algorithm \cite{Mirjalili_Mirjalili_Lewis_2014}
@@
 \section{Problem definition}
+
+% TODO: Odôvodnenie výberu optimalizačného algoritmu
+% TODO: Detailnejsia analýza hlavného algoritmu
+% TODO: Návrh spôsobu nastavenia parametrov
+% TODO: Návrh overenia
 
@@
 
+Majority of datasets available for data science research of transactions fraud detection have the same characteristics - highly imbalanced data, a lot of features and majority of features are anonymized.
+
+Our method will be trained and evaluated on dataset from Kaggle competition EEE-CIS Fraud Detection\footnote{https://www.kaggle.com/c/ieee-fraud-detection/data}.
+As usual in problems of transactions fraud detection, a lot of features of different types are available - 434 features describing demography, credit card, transaction itself, etc.
+Majority of features are anonymized or the meaning of them is not clear.
+In this case, we must be careful to avoid labels leak from some features and also in interpreting data analysis or models result when talking about anonymized features.
+This dataset contains almost 600k samples, that is appropriate for machine learning algorithms.
+However, data are highly imbalanced - classes distribution can bee seen at figure~\ref{fig:classes}.
+When training and evaluating models, one should be careful when data are imbalanced.
+
+\begin{figure}[ht]
+	\begin{center}
+	    \includegraphics[width=0.5\textwidth]{figures/class_distribution.png}
+    \end{center}
+	\caption{Predicted classes distribution - data are highly imbalanced.}
+	\label{fig:classes}
+\end{figure}
+
 \section{Method proposal}
@@
 % TODO: mozeno spomenut oversampling/undersampling
+
+% usage of nature inspired algorithm proved to optimize feature selection much better than baseline well-known methods and overperforms  using power of nature inspired algorithms to address the problem of a lot of anonymized features available in existing datasets.
+Combination of machine learning and nature inspired algorithms proved to be the wa
+
+% Proper data analysis in combination with training machine learning models can help to prevent frauds.
+Problem of choosing correct features without knowing their meaning with goal to improve detection performance can be seen as optimization problem.
+Nature inspired algorithms are known to be very powerful for optimization problems.
 

### 2020-05-01 paper/main.tex
Add preprocessing part to paper

@@
 
-% TODO: Odôvodnenie výberu optimalizačného algoritmu
-% TODO: Detailnejsia analýza hlavného algoritmu
-% TODO: Návrh spôsobu nastavenia parametrov
-% TODO: Návrh overenia
-
 % TODO: opisat nas problem
@@
 
+% TODO: Odôvodnenie výberu optimalizačného algoritmu
+% TODO: Detailnejsia analýza hlavného algoritmu
+% TODO: Návrh spôsobu nastavenia parametrov
+% TODO: Návrh overenia
+
 \section{Preparation}
+
+Before performing experiments, we preprocessed the data.
+At first, useless columns like id of transaction or those with too many missing values (more than 50\%) are dropped.
+Our preprocessing pipeline can be devided into two branches - preprocessing of numeric and categorical attributes.
+In numerical attributes, missing values are filled in with mean value, then all values are normalized (some machine learning algorithms are sensitive to different scales).
+For categorical attributes, transformation of emails to email domains has been performed, missing values have been filled with most frequent values, small categories (smaller than 5\%) merged into one \textit{other} category and at the end, all features have been one-hot encoded.
+Also, missing values indicators have been added to both, numerical and categorical attributes.
 
@@
 % TODO: len kratko spomenut ze preco sme sa rozhodli pouzit aky model - decision tree
+% TODO: model selection phase
 

### 2020-05-27 paper/main.tex
Add abstract to paper

@@
 %
-%\titlerunning{Abbreviated paper title}
+\titlerunning{Transactions Fraud Detection using ML and NIA}
 % If the paper title is too long for the running head, you can set
@@
 \begin{abstract}
-The abstract should briefly summarize the contents of the paper in
-15--250 words.
+Credit card payments are still more and more preferred because of convenience.
+People are using credit cards to pay not only at stores, but also in e-shops.
+However, with the rise of credit cards usage, suitable environment for frauds in transactions arises.
+The data of transactions are usually collected by the banks, that indicates usage of data-mining approaches to prevent frauds.
+The performance of model is based on several sub-steps and decisions, like data analysis, preprocessing, feature selection and hyperparameter optimization.
+In the area of transactions fraud detection, datasets usually contain a lot of features (several hundreds) that are however mostly anonymized.
+Thus we decided to focus on improving model mainly by selecting appropriate features using nature inspired algorithms.
+Nature inspired algorithms proved to be powerful methods in such optimization tasks.
+Experiments showed that proper combination of machine learning model with feature selection using nature inspired algorithms has potential to help to prevent transaction frauds.
 
-\keywords{Transactions Fraud Detection \and Machine Learning \and Nature Inspired Algorithms \and Data Analysis.}
+\keywords{transactions fraud detection \and machine learning classification \and nature inspired algorithms \and feature selection \and bat algorithm.}
 \end{abstract}
@@
 
-% TODO: pribuzne prace na feature selection
-% TODO: spomenut pouzite algoritmy
-% TODO: spomenut domeny, datasety, charakteristky datasetov
-
-% Required: Podobné práce / Related Work - ako riešili vašu alebo podobnú úlohu v
-% existujúcich prácach - nezabudnite v texte uvádzať citácie na práce.
-
 %  opis použitého algoritmu - aj s odôvodnením výberu algoritmu
@@
 % citácii Mendeley)
-
-- Gray Wolf Optimization Algorithm \cite{Mirjalili_Mirjalili_Lewis_2014}
 

### 2020-05-28 src/analysis.py
Add cross-validation for nia search and update notebook

@@ -37,11 +37,51 @@ def show_optimization_run_plot(task, title=''):
     for index, value in enumerate(gen_scores_min):
         data.append([index, value, 'min'])
 
-    df = pd.DataFrame(data, columns=['generation', 'score', 'type'])
+    df = pd.DataFrame(data, columns=['Generation', 'CV score (roc auc)', 'Metric'])
 
     fig, ax = plt.subplots(figsize=(10, 8))
-    ax.set(ylim=(0.5, 1))
-    sns.lineplot(data=df, x='generation', y='score', hue='type', ax=ax).set_title(title)
+    ax.set(ylim=(0.55, 0.85))
+    sns.lineplot(data=df, x='Generation', y='CV score (roc auc)', hue='Metric', ax=ax).set_title(title)
 
     plt.show()
 
+
+def show_optimization_run_plot_for_multiple_tasks(tasks, title='', **kwargs):
+    """
+    Plot the run of multiple optimization tasks by generation.
+    :param task: a tasks to be plotted
+    :param metric: the metric to be plotted
+    :param title: the title of the plot
+    :return: None
+    """
+
+    metric = kwargs.get('metric', 'max')
+    
+    data = []
+    
+    for run, task in enumerate(tasks):
+        # invert the scores because the library did the minimization task
+        gen_scores_max = invert(task.gen_scores_min)
+        gen_scores_min = invert(task.gen_scores_max)
+        gen_scores_mean = invert(task.gen_scores_mean)
+
+        for generation, value in enumerate(gen_scores_mean):
+            data.append([run, generation, value, 'mean'])
+
+        for generation, value in enumerate(gen_scores_max):
+            data.append([run, generation, value, 'max'])
+
+        for generation, value in enumerate(gen_scores_min):
+            data.append([run, generation, value, 'min'])    
+    
+    df = pd.DataFrame(data, columns=['Run', 'Generation', 'CV score (roc auc)', 'Type'])
+    df = df[df['Type'] == metric]
+
+    for ylim in [(0.65, 0.8), None]:
+        fig, ax = plt.subplots(figsize=(10, 8))
+            
+        if ylim is not None:
+            ax.set(ylim=ylim)
+            
+        sns.lineplot(data=df[['Generation', 'CV score (roc auc)', 'Run']], x='Generation', y='CV score (roc auc)', hue='Run', ax=ax).set_title(title)
+        plt.show()

### 2020-05-28 src/classification_benchmark_cv.py

@@ -1,0 +1,100 @@
+from NiaPy.benchmarks import Benchmark
+from sklearn.model_selection import KFold
+from functools import reduce
+
+import numpy as np
+
+
+class ClassificationBenchmarkCV(Benchmark):
+    """
+    NiaPy benchmark for classification task with cross-validation.
+
+    :param model_fn: function which returns sklearn model.
+    :param eval_fn: function(y_test, y_pred) which evaluates predictions
+         and returns a scalar.
+    :param x_train: train data.
+    :param y_train: train labels.
+    :param x_test: test data.
+    :param y_test: test labels.
+    :param cv: number of cross-validation folds.
+    :param random_state: random_state for cross-validation.
+    """
+    def __init__(self, model_fn, eval_fn, x_train, y_train, cv=3, random_state=None):
+        self.x_train = x_train
+        self.y_train = y_train
+        self.model_fn = model_fn
+        self.eval_fn = eval_fn
+        self.k_fold = KFold(n_splits=cv, random_state=random_state, shuffle=True)
+        
+        self.cache = {}
+
+        Benchmark.__init__(self, 0, 1)
+
+    def get_length(self):
+        """
+        Get length of the vector which is being optimized.
+
+        :return: length of the vector which is being optimized.
+        """
+        return len(self.x_train.columns)
+
+    def select_columns(self, solution_vec):
+        """
+        Select columns based on the solution vector.
+
+        :param solution_vec: solution of the problem as a vector.
+        :return: list of column names based on the solution vector.
+        """
+        return self.x_train.columns[solution_vec >= 0.5].tolist()
+    
+    def get_solution_vec_key(self, solution_vec):
+        """
+        Get cache key from the solution vector.
+
+        :param solution_vec: solution of the problem as a vector.
+        :return: cache key string.
+        """
+        return int(reduce(lambda a, b: f"{a}{int(b)}", solution_vec >= 0.5, ""), 2)    
+
+    def function(self):
+        def evaluate(_, solution_vec):
+            solution_vec_key = self.get_solution_vec_key(solution_vec)
+            
+            if solution_vec_key in self.cache:
+                return self.cache[solution_vec_key]
+            
+            selected_columns = self.select_columns(solution_vec)
+
+            # fix of incorrect serialization when using multi threading module
+            if len(selected_columns) == 1 and \
+                    not isinstance(selected_columns[0], str):
+                selected_columns = selected_columns[0]
+
+            if len(selected_columns) < 1:
+                # inverted score, since the optimizer minimizes the task
[…]

### 2020-05-28 src/helpers.py

@@ -1,0 +1,17 @@
+def split_pascal_case_string(input_str):
+    """
+    Split the string which is formatted in "PascalCase" to "Pascal Case"
+    :param input_str: a string to be splitted
+    :return: new string
+    """
+    output_str = ""
+    
+    for index, char in enumerate(input_str):
+        low_char = char.lower()
+
+        if low_char == char or not index:
+            output_str += char
+        else:
+            output_str += f" {char}"
+            
+    return output_str

### 2020-05-29 paper/main.tex
Write experiments and conclusion sections

@@
 % \renewcommand\UrlFont{\color{blue}\rmfamily}
+
+\usepackage{booktabs}
+\usepackage{subfig}
 
@@
 
+We did several experiemnts in which we compared the following algorithms:
+
+\begin{itemize}
+	\item Firefly Algorithm \cite{fister2013comprehensive} - FA (alpha = 1, betamin = 1, gamma = 2)
+	\item Cuckoo Search \cite{yang2009cuckoo} - CS (pa = 0.2, alpha = 0.5)
+	\item Bat Algorithm \cite{yang2010new} - BA (A = 0.5, r = 0.5, Qmin = 0.0, Qmax = 2.0)
+	\item Flower Pollination Algorithm \cite{yang2012flower} - FPA (p = 0.35, beta = 1.5)
+	\item Grey Wolf Optimizer \cite{Mirjalili_Mirjalili_Lewis_2014} - GWO
+\end{itemize}
+
+As a baseline models we have chosen the model with all features and model with selected features from \textit{recursive feature selection with cross-validation}.
+
+\subsection{Experiments strategy}
+
+We chose a \textit{decision tree classifier} as a model on which we will try to find the optimal features. We kept the default hyper-parameters for this classifier set by scikit-learn library.
+
+Since the dataset, if pretty large, firstly we \textit{undersampled} it to achieve the equal size of both classes.
+Then we shrunk it again to 5\% of its size, so we would be able to make as many experiments as possible (we did not have enough computing power).
+We did a train/test split of 80/20.
+
+For each algorithm, we tried several population sizes (10, 20, and 50), for which we did 5 runs of the algorithm.
+We trained exactly 50 generations.
+For both, recursive feature selection and feature selection using nature-inspired algorithms we did 5 fold cross-validation.
+We did not experiment with algorithm parameters and we used their defaults which are described in the previous section.
+
+\subsection{Results}
+
+When comparing the population sizes, all of the algorithms achieved the best results when the population size was highest (50).
+The only exception is Cuckoo Search, which found the most optimal solution with population size 10, however the difference negligible.
+Table \ref{tab:nia_comparison}.
+shows the algorithm scores for each population.
+Figure \ref{fig:nia_train_score_by_algorithm}.
+shows the differences in scores between population size of 10 and 50.
+
+\begin{figure}
+    \centering
+    \subfloat{{ 
+        \includegraphics[width=5.5cm]{figures/nia_train_score_by_algorithm_10}
+    }}
+    \qquad
+    \subfloat{{
+        \includegraphics[width=5.5cm]{figures/nia_train_score_by_algorithm_50}
+    }}
+    \caption{Scores of the algorithms by the population size. The scores are slightly higher for population size of 50.}
+    \label{fig:nia_train_score_by_algorithm}
+\end{figure}
+
+\begin{table}[]
+	\begin{center}
+		\begin{tabular}{|l|rrrrr|}
+			\hline
+			\textbf{Population size \textbackslash Score} & \multicolumn{1}{l}{\textbf{FA}} & \multicolumn{1}{l}{\textbf{CS}} & \multicolumn{1}{l}{\textbf{BA}} & \multicolumn{1}{l}{\textbf{FPA}} & \multicolumn{1}{l|}{\textbf{GWO}} \\ \hline
+			\textbf{10}                                             & 0.753                           & \textbf{0.752}                           & 0.759                           & 0.754                            & 0.776                             \\
+			\textbf{20}                                             & 0.757                           & 0.746                           & 0.755                           & 0.750                            & 0.771                             \\
+			\textbf{50}                                             & \textbf{0.761}                           & 0.751                           & \textbf{0.761}                           & \textbf{0.756}                            & \textbf{0.782}                             \\ \hline
+		\end{tabular}
+	\end{center}
+	\begin{center}
+		\caption{Highest train scores (ROC AUC) of nature inspired algorithms by population size.
+We may notice that Grey Wolf Algorithm (GWO) outperformed all of the other algorithms.} \label{tab:nia_comparison}
+	\end{center}
+\end{table}
+
+All of the nature-inspired algorithms perform roughly the same, except the Grey Wolf Optimizer which outperformed the others (Figure \ref{fig:nia_train_score_by_algorithm}).
+These differences in results (means) between GWO and other algorithms are significant (t-test; $p-value=0.012$).
+Looking for the explanation, we found out that only Grey Wolf Algorithm and Bat Algorithm converged during the optimization run.
+Figure \ref{fig:nia_scores_by_run} shows the optimization of the Firefly Algorithm and Grey Wolf Optimizer, the difference is noticeable.
+If we fine-tuned the parameters of other algorithms, they might converge as well.
+The fact that GWO does not have any parameters (except the population size) was an advantage for it in our experiments.
+
+% https://towardsdatascience.com/hypothesis-testing-in-machine-learning-using-python-a0dc89e169ce
+
+\begin{figure}
+    \centering
+    \subfloat[Firefly Algorithm]{{ 
+        \includegraphics[width=5.5cm]{figures/fa_scores_by_run}
+    }}
+    \qquad
+    \subfloat[Grey Wolf Optimizer]{{
+        \includegraphics[width=5.5cm]{figures/gwo_scores_by_run}
+    }}
+    \caption{Optimization progress for 5 runs of Grey Wolf Optimizer and Firefly Algorithm. The Firefly Algorithm did not really converge.}
+    \label{fig:nia_scores_by_run}
+\end{figure}
+
+The best features, selected by the optimization process, were later used to train the final model on test data.
+These models were compared to a model trained on all features and the model trained on features selected by recursive feature elimination.
+
+The nature-inspired algorithms found a more optimal set of features on the train dataset than recursive feature elimination.
+However, they achieved worse results on the test dataset.
+The only exception is Grey Wolf Optimizer, which achieved the best train and test score.
+The most noticeable difference in the results is the number of selected features by algorithms.
+All of the algorithms (including recursive feature elimination), selected around $~110$ features, however, GWO selected only 53, this is probably the reason of its results.
+These results are shown in Table \ref{tab:nia_baseline_comparison}.
+Figure \ref{fig:nia_selected_features_by_algorithm} shows that GWO selected fewer features in general not only in its best solution.
+
+\begin{table}[]
+	\begin{center}
+		\begin{tabular}{|lrrr|}
+			\hline
+			\textbf{Feature selection type} & \textbf{Train score} & \textbf{Test score} & \textbf{Selected features} \\ \hline
+			All features                    & N/A                  & 0.713               & 240                                  \\
+			Recursive feature elimination   & 0.747                & 0.719               & 112                                  \\
+			Firefly Algorithm               & 0.761                & 0.663               & 114                                  \\
+			Cuckoo Search                   & 0.753                & 0.718               & 106                                  \\
+			Bat Algorithm                   & 0.761                & 0.689               & 125                                  \\
+			Flower Pollination Algorithm    & 0.756                & 0.706               & 113                                  \\
+			\textbf{Grey Wolf Optimizer}             & \textbf{0.782}                & \textbf{0.733}               & \textbf{53}                                   \\ \hline
+		\end{tabular}
+	\end{center}
+	\begin{center}
+		\caption{Comparison of the baseline models with nature-inspired algorithms.
+The train score stands for the best score from the optimization.
+All of the nature-inspired algorithms (except Grey Wolf Optimizer) achieved a better train score than recursive feature elimination, however, they overfit the model which resulted in a worse test scores.} \label{tab:nia_baseline_comparison}
+	\end{center}
+\end{table}
+
+\begin{figure}[ht]
+	\begin{center}
+	    \includegraphics[width=5.5cm]{figures/nia_selected_features_by_algorithm.png}
+    \end{center}
+	\caption{Number of selected features by algorithm - Grey Wolf Algorithm selected by far the least number of features.}
+	\label{fig:nia_selected_features_by_algorithm}
+\end{figure}
+
 % TODO: ake experiemnty sme spravili
@@
 
+In our experiments, we found out Grey Wolf Optimizer as the best nature-inspired algorithm for feature selection.
+It achieved significantly better results than other nature-inspired algorithms and better results that recursive feature elimination.
+
+GWO alongside Bat Algorithm was the only nature-inspired algorithms that were able to converge during the optimization process.
+We assume that other algorithms might converge too if we experimented with their parameter settings.
+We consider the fact, that GWO does not have any parameters other than population size as a big strength, because it saves a significant amount of optimization time.
+
+In comparison with traditional ways of feature selection (recursive feature elimination), all of the nature-inspired algorithms were able to find better solutions on the training dataset, however, they overfit most of the time.
+Moreover, the feature selection using nature-inspired algorithms was much slower.
+
 \bibliographystyle{splncs04}

### 2020-05-29 paper/main.tex
Update TODOs
Fix mistakes

@@
 
-% TODO: len kratko, co navrhujeme - feature selection na nejaky model pomocou nejakeho prirodou inspirovaneho algoritmu
-% TODO: mozeno spomenut oversampling/undersampling
+% TODO: Len kratko, co navrhujeme - feature selection na nejaky model pomocou nejakeho prirodou inspirovaneho algoritmu
 
@@
 
-Since the dataset, if pretty large, firstly we \textit{undersampled} it to achieve the equal size of both classes.
-Then we shrunk it again to 5\% of its size, so we would be able to make as many experiments as possible (we did not have enough computing power).
+Since the dataset is pretty large, we firstly \textit{undersampled} it to achieve the equal size of both classes.
+Then we shrunk it again to 5\% of its size, so we would be able to make as many experiments as possible (since we did not have enough computing power).
 We did a train/test split of 80/20.
 
-For each algorithm, we tried several population sizes (10, 20, and 50), for which we did 5 runs of the algorithm.
+For each algorithm, we tried several population sizes (10, 20, and 50), for which we did 5 runs of each algorithm.
 We trained exactly 50 generations.
 For both, recursive feature selection and feature selection using nature-inspired algorithms we did 5 fold cross-validation.
-We did not experiment with algorithm parameters and we used their defaults which are described in the previous section.
+We did not experiment with algorithm parameters and we used their defaults, as we mentioned in the previous section.
 
@@
 When comparing the population sizes, all of the algorithms achieved the best results when the population size was highest (50).
-The only exception is Cuckoo Search, which found the most optimal solution with population size 10, however the difference negligible.
+The only exception is Cuckoo Search, which found the most optimal solution with population size of 10, however the difference negligible.
 Table \ref{tab:nia_comparison}.
@@
 All of the nature-inspired algorithms perform roughly the same, except the Grey Wolf Optimizer which outperformed the others (Figure \ref{fig:nia_train_score_by_algorithm}).
-These differences in results (means) between GWO and other algorithms are significant (t-test; $p-value=0.012$).
+These differences in results between GWO and other algorithms are significant (there is a sigificant difference in means, t-test; $p-value=0.012$).
 Looking for the explanation, we found out that only Grey Wolf Algorithm and Bat Algorithm converged during the optimization run.
-Figure \ref{fig:nia_scores_by_run} shows the optimization of the Firefly Algorithm and Grey Wolf Optimizer, the difference is noticeable.
+Figure \ref{fig:nia_scores_by_run} shows the optimization progress of the Firefly Algorithm and Grey Wolf Optimizer, Firefly Algorithm did not converge.
 If we fine-tuned the parameters of other algorithms, they might converge as well.
-The fact that GWO does not have any parameters (except the population size) was an advantage for it in our experiments.
+The fact that GWO does not have any parameters (except the population size) might be an advantage for it in our experiments.
 
@@
 
-The best features, selected by the optimization process, were later used to train the final model on test data.
-These models were compared to a model trained on all features and the model trained on features selected by recursive feature elimination.
+The best features, selected in the optimization process, were later used to train the final model on test data.
+These models were compared to the model trained on all features and the model trained on features selected by recursive feature elimination.
 
@@
 
-% TODO: ake experiemnty sme spravili
-% TODO: tabulka porovnania algoritmov
-% TODO: learning along the way with consequences - napr undersampling/oversampling atd.
-
 \section{Conclusion}
-
-% TODO: zhrnutie vysledkov, mozne vylepsenia a uskalia nasho pristupu
 
@@
 
+% TODO: concusion in the topic of transaction fraud detection
+
 \bibliographystyle{splncs04}

### 2020-05-29 paper/main.tex
Add method proposal

@@
 \usepackage{subfig}
+\usepackage{svg}
 
@@
 
-% TODO: Len kratko, co navrhujeme - feature selection na nejaky model pomocou nejakeho prirodou inspirovaneho algoritmu
+% TODO: Len kratko, co navrhujeme - feature selection na nejaky model pomocou nejakeho prirodou inspirovaneho algoritmu.
 
@@
 Nature inspired algorithms are known to be very powerful for optimization problems.
+
+We propose to use nature-inspired algorithms for feature selection.
+We will try several nature-inspired algorithms and compare them to find out which one is most suitable for our problem.
+We decided to use the following algorithms: Firefly Algorithm (FA) \cite{fister2013comprehensive}, Cuckoo Search (CS) \cite{yang2009cuckoo}, Bat Algorithm (BA) \cite{yang2010new}, Flower Pollination Algorithm (FPA) \cite{yang2012flower}, Grey Wolf Optimizer (GWO) \cite{Mirjalili_Mirjalili_Lewis_2014} since they were already used for feature selection in related works or they performed extraordinarily in other types of tasks.
+
+These algorithms will be trying to find the optimal set of features, which will be represented as a binary vector.
+The size of this vector will be equal to the number of features.
+Ones in the vector will indicate that the respective features are selected and nulls vice-versa.
+The nature-inspired algorithms won't be directly optimizing that binary vector, they will be optimizing vector of floats with values in $<0, 1>$ which will be later transformed to binary vector (Figure \ref{fig:projection}).
+The model will be fitted with the selected features from the solution vector.
+
+\begin{figure}[ht]
+	\begin{center}
+	    \includegraphics[width=6cm]{figures/projection.png}
+    \end{center}
+	\caption{Projection from the vector of floats which is being optimized by nature-inspired algorithms to the binary vector representing wether feature is selected or not.}
+	\label{fig:projection}
+\end{figure}
+
+To prevent overfitting we will do cross-validation ($k=5$) for each solution, the mean score from every one of $k$ fits will represent the fitness of the solution.
+
+% TODO: Mention which were previosly used for feature selection
 
@@
 
-We did several experiemnts in which we compared the following algorithms:
+We did several experiemnts in which we compared the following algorithms with these parameters:
 
 \begin{itemize}
-	\item Firefly Algorithm \cite{fister2013comprehensive} - FA (alpha = 1, betamin = 1, gamma = 2)
-	\item Cuckoo Search \cite{yang2009cuckoo} - CS (pa = 0.2, alpha = 0.5)
-	\item Bat Algorithm \cite{yang2010new} - BA (A = 0.5, r = 0.5, Qmin = 0.0, Qmax = 2.0)
-	\item Flower Pollination Algorithm \cite{yang2012flower} - FPA (p = 0.35, beta = 1.5)
-	\item Grey Wolf Optimizer \cite{Mirjalili_Mirjalili_Lewis_2014} - GWO
+	\item Firefly Algorithm (alpha = 1, betamin = 1, gamma = 2)
+	\item Cuckoo Search (pa = 0.2, alpha = 0.5)
+	\item Bat Algorithm (A = 0.5, r = 0.5, Qmin = 0.0, Qmax = 2.0)
+	\item Flower Pollination Algorithm (p = 0.35, beta = 1.5)
+	\item Grey Wolf Optimizer
 \end{itemize}

### 2020-05-30 paper/main.tex
Add related works and extend introduction

@@
 Therefore, transactions fraud detection can be interpreted as data mining problem, concretely binary classification.
+Majority of existing datasets of transactions frauds contain a lot of features, however most of them are usually anonymized.
+Thus, model will rely also on proper feature selection.
+In such cases, where feature selection is very important part of the problem, powerful metaheuristic search methods like nature inspired algorithms can be used.
 
@@
 
-% Required: Úvod / Introduction - aká úloha sa ide riešiť, prečo je dobré ju riešiť
+\section{Related works}
 
-% TODO: co robime, tj, nacrtnut problem atd
-% TODO: spomenut, datasety s velkym mnozstvom features, je potrebna feature selection
-% TODO: spomenut, ze robi sa to takto takto, ale su aj nejako prirodou inspirovane algoritmy, ktore maju taketo vyhody
-% TODO: preco? napr. mzoe chybat domenovat znalost (features) kvoli anonymizacii - nas pripad
+Number of transaction frauds is arising alongside with usage of credit cards.
+To help prevent transaction frauds, common data-mining techniques are available.
+However, data are in most cases very specific and there is no method suitable to all problems.
+Addressing that, comparative study of various machine learning techniques, mostly neural networks, for credit card frauds detection was performed~\cite{Sadgali2019}.
+Common algorithms like Support Vector Machines~\cite{Wiese2009} or Random Forest~\cite{Xuan2018} were also experimented.
+Another and very specific way to detect frauds is using Genetic programming~\cite{Assis2013}.
+Genetic programming has proven to be very effective for this problem, achieving significant improvements (up to almost 17\% compared to corporation baseline).
 
-\section{Related works}
+Each transaction can be viewed separately as one unit, but also whole transactions history can be considered.
+To create model that is resistant to changes in transactions, modeling time series using Long-Short Term Memory (LSTM) recurrent neural networks seems to be the right way~\cite{Wiese2009}.
+
+Publicly available datasets for transactions fraud detection usually contain a lof of features.
+It is necessary to put special effort on feature selection phase in such cases.
+Nature inspired algorithms proved to be very powerful for feature selection problems in previous research.
+When using metaheuristic approaches like nature inspired algorithms, feature selection problem can be interpreted as combinatorial optimization problem~\cite{Wang2016}.
+
+Rough sets are popular when using nature inspired algorithms for feature selection (more about rough sets theory used in data mining in~\cite{Slimani2013}).
+Rough sets were used in combination with Firefly Algorithm~\cite{Banati2011}, Particle Swarm Optimization~\cite{Wang2007} or Bat Algorithm~\cite{Emary2014}.
+Majority of works used multiple datasets for their experiments to prove nature inspired algorithms power.
+
+There are also another surveys focused mainly on usage of nature inspired algorithms for feature selection~\cite{Kauser2018,Wang2016,Yang2020}.
+Further algorithms were researched researched, like Artificial Bees Colony, Cuckoo Search, Ant Colony Optimization or Genetic Algorithms.
 
@@
 Nature inspired algorithms are known to be very powerful for optimization problems.
+
+% 440 features at the end
 

### 2020-05-30 paper/main.tex
Reformat figures and tables

@@
 
+\def\arraystretch{1.5}
 \begin{table}[]
 	\begin{center}
-		\begin{tabular}{|l|rrrrr|}
+		\begin{tabular}{|l|r|r|r|r|r|}
 			\hline
-			\textbf{Population size \textbackslash Score} & \multicolumn{1}{l}{\textbf{FA}} & \multicolumn{1}{l}{\textbf{CS}} & \multicolumn{1}{l}{\textbf{BA}} & \multicolumn{1}{l}{\textbf{FPA}} & \multicolumn{1}{l|}{\textbf{GWO}} \\ \hline
+			\textbf{Population size \textbackslash Score} & \multicolumn{1}{l|}{\textbf{FA}} & \multicolumn{1}{l|}{\textbf{CS}} & \multicolumn{1}{l|}{\textbf{BA}} & \multicolumn{1}{l|}{\textbf{FPA}} & \multicolumn{1}{l|}{\textbf{GWO}} \\ \hline
 			\textbf{10}                                             & 0.753                           & \textbf{0.752}                           & 0.759                           & 0.754                            & 0.776                             \\
@@
 	\begin{center}
-		\begin{tabular}{|lrrr|}
+		\begin{tabular}{|l|r|r|r|}
 			\hline
@@
 	\begin{center}
-	    \includegraphics[width=5.5cm]{figures/nia_selected_features_by_algorithm.png}
+	    \includegraphics[width=7.5cm]{figures/nia_selected_features_by_algorithm.png}
     \end{center}

### 2020-05-31 paper/main.tex
Add finishing touches

--- 2020-05-29 paper/main.tex
@@
 %
-\title{Transactions Fraud Detection using Machine Learning and Nature Inspired Algorithms}
+\title{Using Nature Inspired Algorithms for Feature Selection in Transaction Fraud Detection}
 %
-\titlerunning{Transactions Fraud Detection using ML and NIA}
+\titlerunning{Using NIA for Feature Selection in Transaction Fraud Detection}
 % If the paper title is too long for the running head, you can set
@@
 Credit card payments are still more and more preferred because of convenience.
-People are using credit cards to pay not only at stores, but also in e-shops.
-However, with the rise of credit cards usage, suitable environment for frauds in transactions arises.
-The data of transactions are usually collected by the banks, that indicates usage of data-mining approaches to prevent frauds.
-The performance of model is based on several sub-steps and decisions, like data analysis, preprocessing, feature selection and hyperparameter optimization.
-In the area of transactions fraud detection, datasets usually contain a lot of features (several hundreds) that are however mostly anonymized.
-Thus we decided to focus on improving model mainly by selecting appropriate features using nature inspired algorithms.
-Nature inspired algorithms proved to be powerful methods in such optimization tasks.
-Experiments showed that proper combination of machine learning model with feature selection using nature inspired algorithms has potential to help to prevent transaction frauds.
+People are using credit cards to pay not only at stores but also in e-shops.
+However, with the rise of credit card usage, a suitable environment for fraud in transactions arises.
+The data of transactions are usually collected by the banks, which indicates the usage of data-mining approaches to prevent frauds.
+The performance of the model is based on several sub-steps and decisions, like data analysis, preprocessing, feature selection, and hyperparameter optimization.
+In the area of transaction fraud detection, datasets usually contain a lot of features (several hundred) that are however mostly anonymized.
+Thus we decided to focus on improving model mainly by selecting appropriate features using nature-inspired algorithms.
+Nature-inspired algorithms proved to be powerful methods in such optimization tasks.
+Experiments showed that a proper combination of machine learning model with feature selection using nature-inspired algorithms has the potential to help to prevent transaction frauds.
+Out of 5 nature-inspired algorithms, the Grey Wolf Optimizer achieved the best results.
+It improved the model by 1.4\% (ROC AUC) while selecting half of the features in comparison with a recursive feature elimination.
 
-\keywords{transactions fraud detection \and machine learning classification \and nature inspired algorithms \and feature selection \and bat algorithm.}
+\keywords{transaction fraud detection \and machine learning classification \and nature inspired algorithms \and feature selection \and bat algorithm.}
 \end{abstract}
@@
 Payments with credit cards are nowadays still more preferred over using cash.
-Using credit card is not only more comfortable for people, but even also more safe than carrying cash in wallet when it comes to higher amount of money.
-However, number of transaction frauds is arising alongside with usage of credit cards.
-Number of transactions and ability to obtain data from them indicate the need of automatic detection of suspicious payments.
+Using a credit card is not only more comfortable for people, but even also more safe than carrying cash in wallet when it comes to a higher amount of money.
+However, the number of transaction frauds is arising alongside with usage of credit cards.
+The number of transactions and the ability to obtain data from them indicates the need for automatic detection of suspicious payments.
 
 Data of performed transactions via credit cards are naturally collected by companies and banks to produce statistics or investigate frauds.
-Therefore, transactions fraud detection can be interpreted as data mining problem, concretely binary classification.
+Therefore, transaction fraud detection can be interpreted as a data mining problem, concretely binary classification.
+The majority of existing datasets of transaction frauds contain a lot of features, however, most of them are usually anonymized.
+Thus, the model will rely also on proper feature selection.
+In such cases, where feature selection is a very important part of the problem, powerful metaheuristic search methods like nature-inspired algorithms can be used.
 
-In this paper, we propose novel method for detection frauds in transactions using aspects of data analysis, machine learning and nature inspired algorithms.
-The basis of our method lies in training machine learning model on best features selected by nature inspired algorithm.
-More nature inspired algorithms are compared to choose the best one for this problem.
-Finally, nature inspired algorithms in combination with machine learning proved to be very efficient way for detecting frauds in transactions, overperforming common methods in this area.
-% TODO ten overperforming treba nejak doladit mozno
-
-% Required: Úvod / Introduction - aká úloha sa ide riešiť, prečo je dobré ju riešiť
-
-% TODO: co robime, tj, nacrtnut problem atd
-% TODO: spomenut, datasety s velkym mnozstvom features, je potrebna feature selection
-% TODO: spomenut, ze robi sa to takto takto, ale su aj nejako prirodou inspirovane algoritmy, ktore maju taketo vyhody
-% TODO: preco? napr. mzoe chybat domenovat znalost (features) kvoli anonymizacii - nas pripad
+In this paper, we propose a novel method for detection frauds in transactions using aspects of data analysis, machine learning, and nature-inspired algorithms.
+The basis of our method lies in the training machine learning model on the best features selected by the nature-inspired algorithm.
+More nature-inspired algorithms are compared to choose the best one for this problem.
 
 \section{Related works}
+
+The number of transaction frauds is arising alongside with usage of credit cards.
+To help prevent transaction frauds, common data-mining techniques are available.
+However, data are in most cases very specific and there is no method suitable to all problems.
+Addressing that, a comparative study of various machine learning techniques, mostly neural networks, for credit card fraud detection was performed~\cite{Sadgali2019}.
+Common algorithms like Support Vector Machines~\cite{Wiese2009} or Random Forest~\cite{Xuan2018} were also examined.
+Another and very specific way to detect frauds is by using Genetic programming~\cite{Assis2013}.
+Genetic programming has proven to be very effective for this problem, achieving significant improvements (up to almost 17\% compared to corporation baseline).
+
+Each transaction can be viewed separately as one unit, but also whole transaction history can be considered.
+To create a model that is resistant to changes in transactions, modeling time series using Long-Short Term Memory (LSTM) recurrent neural networks seems to be the right way~\cite{Wiese2009}.
+
+Publicly available datasets for transaction fraud detection usually contain a lot of features.
+It is necessary to put special effort into the feature selection phase in such cases.
+Nature-inspired algorithms proved to be very powerful for feature selection problems in previous research.
+When using metaheuristic approaches like nature-inspired algorithms, feature selection problem can be interpreted as combinatorial optimization problem~\cite{Wang2016}.
+
+Rough sets are popular when using nature-inspired algorithms for feature selection (more about rough sets theory used in data mining in~\cite{Slimani2013}).
+Rough sets were used in combination with Firefly Algorithm~\cite{Banati2011}, Particle Swarm Optimization~\cite{Wang2007}, or Bat Algorithm~\cite{Emary2014}.
+The majority of works used multiple datasets for their experiments to prove nature-inspired algorithms power.
+
+There are also other surveys focused mainly on the usage of nature-inspired algorithms for feature selection~\cite{Kauser2018,Wang2016,Yang2020}.
+Further algorithms were researched, like Artificial Bees Colony, Cuckoo Search, Ant Colony Optimization, or Genetic Algorithms.
 
@@
 
-Majority of datasets available for data science research of transactions fraud detection have the same characteristics - highly imbalanced data, a lot of features and majority of features are anonymized.
+The majority of datasets available for data science research of transaction fraud detection have the same characteristics - highly imbalanced data, a lot of features, and the majority of features are anonymized.
 
-Our method will be trained and evaluated on dataset from Kaggle competition EEE-CIS Fraud Detection\footnote{https://www.kaggle.com/c/ieee-fraud-detection/data}.
-As usual in problems of transactions fraud detection, a lot of features of different types are available - 434 features describing demography, credit card, transaction itself, etc.
-Majority of features are anonymized or the meaning of them is not clear.
+Our method will be trained and evaluated on a dataset from Kaggle competition EEE-CIS Fraud Detection\footnote{https://www.kaggle.com/c/ieee-fraud-detection/data}.
+As usual in problems of transaction fraud detection, a lot of features of different types are available - 434 features describing demography, credit card, the transaction itself, etc.
+The majority of features are anonymized or the meaning of them is not clear.
 In this case, we must be careful to avoid labels leak from some features and also in interpreting data analysis or models result when talking about anonymized features.
 This dataset contains almost 600k samples, that is appropriate for machine learning algorithms.
-However, data are highly imbalanced - classes distribution can bee seen at figure~\ref{fig:classes}.
+However, data are highly imbalanced - class distribution can be seen in figure~\ref{fig:classes}.
 When training and evaluating models, one should be careful when data are imbalanced.
+
+We cannot handpick the features, because the data are anonymized so we lack the domain knowledge.
+Hence we need to apply some methods for feature selection.
+Also, manually hand-picking for that large amount of features would be nearly impossible.
+Another motivation for feature selection is that it reduces model complexity and improves its interpretability.
 
@@
 
-We propose to use nature-inspired algorithms for feature selection.
+% 440 features at the end
+
+To improve a machine learning models for the transaction fraud detection in a way of interpretability and complexity, we propose to use nature-inspired algorithms for feature selection.
 We will try several nature-inspired algorithms and compare them to find out which one is most suitable for our problem.
@@
 	\begin{center}
-	    \includegraphics[width=6cm]{figures/projection.png}
+	    \includegraphics[width=8cm]{figures/projection.png}
     \end{center}
@@
 
-We did several experiemnts in which we compared the following algorithms with these parameters:
+We did several experiemnts in which we compared the following algorithms and their parameters:
 
 \begin{itemize}
-	\item Firefly Algorithm (alpha = 1, betamin = 1, gamma = 2)
-	\item Cuckoo Search (pa = 0.2, alpha = 0.5)
-	\item Bat Algorithm (A = 0.5, r = 0.5, Qmin = 0.0, Qmax = 2.0)
-	\item Flower Pollination Algorithm (p = 0.35, beta = 1.5)
+	\item Firefly Algorithm ($\alpha = 1$, $\beta_{min} = 1$, $\gamma = 2$)
+	\item Cuckoo Search ($p_a = 0.2$, $\alpha = 0.5$)
+	\item Bat Algorithm ($A = 0.5$, $r = 0.5$, $Q_{min} = 0.0$, $Q_{max} = 2.0$)
+	\item Flower Pollination Algorithm ($p = 0.35$, $\beta = 1.5$)
 	\item Grey Wolf Optimizer
@@
 
-As a baseline models we have chosen the model with all features and model with selected features from \textit{recursive feature selection with cross-validation}.
+We will use the algorithm default parameters as they are implemented in \textit{NiaPy} \cite{NiaPyJOSS2018} library based on the original papers.
+As a baseline models we have chosen the model with all features and model with selected features from \textit{recursive feature elimination with cross-validation}.
 
@@
 
-We chose a \textit{decision tree classifier} as a model on which we will try to find the optimal features. We kept the default hyper-parameters for this classifier set by scikit-learn library.
+We chose a \textit{decision tree classifier} as a model on which we will try to find the optimal features.
+Based on our experiments it achieved the best results considering also the training speed.
+We kept the default hyper-parameters for this classifier set by scikit-learn library.
 
@@
 We trained exactly 50 generations.
-For both, recursive feature selection and feature selection using nature-inspired algorithms we did 5 fold cross-validation.
-We did not experiment with algorithm parameters and we used their defaults, as we mentioned in the previous section.
+For both, recursive feature elimination and feature selection using nature-inspired algorithms we did 5 fold cross-validation.
 
@@
 shows the algorithm scores for each population.
-Figure \ref{fig:nia_train_score_by_algorithm}.
-shows the differences in scores between population size of 10 and 50.
+Figure \ref{fig:nia_train_score_by_algorithm_1}.
+\& \ref{fig:nia_train_score_by_algorithm_2}.shows the differences in scores between population size of 10 and 50.
+
+% \begin{figure}
+%     \centering
+%     \subfloat{{ 
+%         \includegraphics[width=5.5cm]{figures/nia_train_score_by_algorithm_10}
+%     }}
+%     \qquad
+%     \subfloat{{
+%         \includegraphics[width=5.5cm]{figures/nia_train_score_by_algorithm_50}
+%     }}
+%     \caption{Scores of the algorithms by the population size. The scores are slightly higher for population size of 50.}
+%     \label{fig:nia_train_score_by_algorithm}
+% \end{figure}
 
@@
     \centering
-    \subfloat{{ 
-        \includegraphics[width=5.5cm]{figures/nia_train_score_by_algorithm_10}
-    }}
-    \qquad
-    \subfloat{{
-        \includegraphics[width=5.5cm]{figures/nia_train_score_by_algorithm_50}
-    }}
-    \caption{Scores of the algorithms by the population size. The scores are slightly higher for population size of 50.}
-    \label{fig:nia_train_score_by_algorithm}
+    \includegraphics[width=11cm]{figures/nia_train_score_by_algorithm_10}
+    \caption{Scores of the nature-inspired algorithms for population size of 10.}
+    \label{fig:nia_train_score_by_algorithm_1}
 \end{figure}
 
-\begin{table}[]
+\begin{figure}
+    \centering
+    \includegraphics[width=11cm]{figures/nia_train_score_by_algorithm_10}
+    \caption{Scores of the nature-inspired algorithms for population size of 50.}
+    \label{fig:nia_train_score_by_algorithm_2}
+\end{figure}
+
+\def\arraystretch{1.5}
+\begin{table}
 	\begin{center}
-		\begin{tabular}{|l|rrrrr|}
+		\caption{Highest train scores (ROC AUC) of nature inspired algorithms by population size.
+We may notice that Grey Wolf Algorithm (GWO) outperformed all of the other algorithms.} \label{tab:nia_comparison}
+	\end{center}
+	\begin{center}
+		\begin{tabular}{|l|r|r|r|r|r|}
 			\hline
-			\textbf{Population size \textbackslash Score} & \multicolumn{1}{l}{\textbf{FA}} & \multicolumn{1}{l}{\textbf{CS}} & \multicolumn{1}{l}{\textbf{BA}} & \multicolumn{1}{l}{\textbf{FPA}} & \multicolumn{1}{l|}{\textbf{GWO}} \\ \hline
+			\textbf{Population size \textbackslash Score} & \multicolumn{1}{r|}{\textbf{FA}} & \multicolumn{1}{r|}{\textbf{CS}} & \multicolumn{1}{r|}{\textbf{BA}} & \multicolumn{1}{r|}{\textbf{FPA}} & \multicolumn{1}{r|}{\textbf{GWO}} \\ \hline
 			\textbf{10}                                             & 0.753                           & \textbf{0.752}                           & 0.759                           & 0.754                            & 0.776                             \\
@@
 		\end{tabular}
-	\end{center}
-	\begin{center}
-		\caption{Highest train scores (ROC AUC) of nature inspired algorithms by population size.
-We may notice that Grey Wolf Algorithm (GWO) outperformed all of the other algorithms.} \label{tab:nia_comparison}
 	\end{center}
@@
 Looking for the explanation, we found out that only Grey Wolf Algorithm and Bat Algorithm converged during the optimization run.
-Figure \ref{fig:nia_scores_by_run} shows the optimization progress of the Firefly Algorithm and Grey Wolf Optimizer, Firefly Algorithm did not converge.
+Figure \ref{fig:nia_scores_by_run_1}.
+\& \ref{fig:nia_scores_by_run_2}.
+shows the optimization progress of the Firefly Algorithm and Grey Wolf Optimizer, Firefly Algorithm did not converge.
 If we fine-tuned the parameters of other algorithms, they might converge as well.
@@
 
-\begin{figure}
+\begin{figure}[htp]
 	\centering
-    \subfloat[Firefly Algorithm]{{ 
-        \includegraphics[width=5.5cm]{figures/fa_scores_by_run}
-    }}
-    \qquad
-    \subfloat[Grey Wolf Optimizer]{{
-        \includegraphics[width=5.5cm]{figures/gwo_scores_by_run}
-    }}
-    \caption{Optimization progress for 5 runs of Grey Wolf Optimizer and Firefly Algorithm. The Firefly Algorithm did not really converge.}
-    \label{fig:nia_scores_by_run}
+	\includegraphics[clip, width=10.5cm]{figures/gwo_scores_by_run}
+    \caption{Optimization progress for 5 runs of Grey Wolf Optimizer.}
+    \label{fig:nia_scores_by_run_1}
+\end{figure}
+
+\begin{figure}[htp]
+	\centering
+	\includegraphics[clip, width=10.5cm]{figures/fa_scores_by_run}
+    \caption{Optimization progress for 5 runs of Firefly Algorithm. The optimization process did not converge.}
+    \label{fig:nia_scores_by_run_2}
 \end{figure}
@@
 	\begin{center}
-		\begin{tabular}{|lrrr|}
+		\caption{Comparison of the baseline models with nature-inspired algorithms.
+The train score stands for the best score from the optimization.
+All of the nature-inspired algorithms (except Grey Wolf Optimizer) achieved a better train score than recursive feature elimination, however, they overfit the model which resulted in a worse test scores.} \label{tab:nia_baseline_comparison}
+	\end{center}
+	\begin{center}
+		\begin{tabular}{|l|r|r|r|}
 			\hline
@@
 	\end{center}
-	\begin{center}
-		\caption{Comparison of the baseline models with nature-inspired algorithms.
-The train score stands for the best score from the optimization.
-All of the nature-inspired algorithms (except Grey Wolf Optimizer) achieved a better train score than recursive feature elimination, however, they overfit the model which resulted in a worse test scores.} \label{tab:nia_baseline_comparison}
-	\end{center}
 \end{table}
@@
 	\begin{center}
-	    \includegraphics[width=5.5cm]{figures/nia_selected_features_by_algorithm.png}
+	    \includegraphics[width=10cm]{figures/nia_selected_features_by_algorithm.png}
     \end{center}

### 2020-05-31 paper/main.tex
Update conclusion

@@
 In our experiments, we found out Grey Wolf Optimizer as the best nature-inspired algorithm for feature selection.
-It achieved significantly better results than other nature-inspired algorithms and better results that recursive feature elimination.
+It achieved significantly better results than other nature-inspired algorithms and better results (1.4\% ROC AUC) that recursive feature elimination, while using by half fewer features.
 

### 2020-05-31 paper/main.tex
Fix alignment of the document

--- 2020-05-31 paper/main.tex
@@
 Figure \ref{fig:nia_train_score_by_algorithm_1}.
-\& \ref{fig:nia_train_score_by_algorithm_2}.shows the differences in scores between population size of 10 and 50.
+\& \ref{fig:nia_train_score_by_algorithm_2}.
+shows the differences in scores between population size of 10 and 50.
 
@@
     \includegraphics[width=11cm]{figures/nia_train_score_by_algorithm_10}
-    \caption{Scores of the nature-inspired algorithms for population size of 10.}
+    \caption{Scores of the nature-inspired in 5 runs algorithms for population size of 10.}
     \label{fig:nia_train_score_by_algorithm_1}
@@
     \includegraphics[width=11cm]{figures/nia_train_score_by_algorithm_10}
-    \caption{Scores of the nature-inspired algorithms for population size of 50.}
+    \caption{Scores of the nature-inspired algorithms in 5 runs for population size of 50.}
     \label{fig:nia_train_score_by_algorithm_2}
@@
 
-\begin{figure}[ht]
+\begin{figure}[h!]
 	\begin{center}
@@
     \end{center}
-	\caption{Number of selected features by algorithm - Grey Wolf Algorithm selected by far the least number of features.}
+	\caption{Number of selected features by algorithm in 5 runs - Grey Wolf Algorithm selected by far the least number of features.}
 	\label{fig:nia_selected_features_by_algorithm}
 \end{figure}
+
+\newpage
 
@@
 In our experiments, we found out Grey Wolf Optimizer as the best nature-inspired algorithm for feature selection.
-It achieved significantly better results than other nature-inspired algorithms and better results that recursive feature elimination.
+It achieved significantly better results than other nature-inspired algorithms.
+Grey Wolf Optimizer achieved even better results than recursive feature elimination (+1.4\% ROC AUC), while using by half fewer features.
 
-GWO alongside Bat Algorithm was the only nature-inspired algorithms that were able to converge during the optimization process.
+Grey Wolf Optimizer alongside Bat Algorithm was the only nature-inspired algorithms that were able to converge during the optimization process.
 We assume that other algorithms might converge too if we experimented with their parameter settings.
@@
 
-% TODO: concusion in the topic of transaction fraud detection
-
 \bibliographystyle{splncs04}

### 2020-05-31 paper/main.tex
Fix figure
Fix details in paper

@@
     \includegraphics[width=11cm]{figures/nia_train_score_by_algorithm_10}
-    \caption{Scores of the nature-inspired in 5 runs algorithms for population size of 10.}
+    \caption{Scores of the nature-inspired algorithms in 5 runs - population size of 10.}
     \label{fig:nia_train_score_by_algorithm_1}
@@
     \centering
-    \includegraphics[width=11cm]{figures/nia_train_score_by_algorithm_10}
-    \caption{Scores of the nature-inspired algorithms in 5 runs for population size of 50.}
+    \includegraphics[width=11cm]{figures/nia_train_score_by_algorithm_50}
+    \caption{Scores of the nature-inspired algorithms in 5 runs - population size of 50.}
     \label{fig:nia_train_score_by_algorithm_2}
@@
 \begin{table}
+	\vspace*{0.5cm}
 	\begin{center}
@@
 
-All of the nature-inspired algorithms perform roughly the same, except the Grey Wolf Optimizer which outperformed the others (Figure \ref{fig:nia_train_score_by_algorithm}).
+All of the nature-inspired algorithms perform roughly the same, except the Grey Wolf Optimizer which outperformed the others (Figure \ref{fig:nia_train_score_by_algorithm_1} \& \ref{fig:nia_train_score_by_algorithm_2}).
 These differences in results between GWO and other algorithms are significant (there is a sigificant difference in means, t-test; $p-value=0.012$).
@@
 \begin{figure}[h!]
+	\vspace*{1cm}
 	\begin{center}
